The Rockerverse: Packages and Applications for Containerisation with R

### 2020-03-04 rockerverse.Rmd
add teaching use case

**Teaching** is a further example of where shared computing environments and sandboxing can greatly improve the process.
First, _prepared environments for teaching_ are especially helpful for (a) introductory courses, where students often struggle with the first step of installation and configuration \citep{cetinkaya-rundel_infrastructure_2018}, and (b) courses that require access to a relatively complex setup of software tools, e.g., database systems.
\citet{cetinkaya-rundel_infrastructure_2018} describe how a Docker-based deployment of RStudio (i) avoided problems with troubleshooting individual students' computers and greatly increased engagement through very quickly showing tangible outcomes, e.g., a visualisation, and (ii) reduced demand on teaching and IT staff.
Each student received access to a personal RStudio instance running in a container after authentication with the university login, which gives the benefits of sandboxing and the possibility of limiting resources.
\citet{cetinkaya-rundel_infrastructure_2018} found that for the courses at hand, actual usage of the UI is intermittent so a single cloud-based VM with four cores and 28&nbsp;GB RAM sufficed for over 100 containers.
An example for mitigating _complex setups_ is teaching databases.
R is very useful tool for interfacing with databases, because almost every open source and proprietary database system has an R package that allows users to connect and interact with it.
This flexibility is even broadened by \CRANpkg{DBI} \citep{cran_dbi}, which allows for creating a common API for interfacing these databases, or the \CRANpkg{dbplyr} package \citep{cran_dbplyr}, which runs \CRANpkg{dplyr} \citep{cran_dplyr} code straight against the database as queries.
But learning and teaching these tools comes with the cost of deploying or having access to an environment with the software and drivers installed.
For people teaching R, it can become a barrier if they need to install local versions of database drivers or connect to remote instances which might or might not be made available by IT services.
Giving access to a sandbox for the most common environments for teaching databases is the idea behind [`r-db`](https://github.com/ColinFay/r-db), a Docker image that contains everything needed to connect to a database from R.
Notably, with `r-db`, users do not have to install complex drivers or configure their machine in a specific way.
The `rocker/tidyverse` base image ensures that users can also readily use packages for analysis, display, and reporting.

Second, the idea of a common environment and partitioning allows for using _containers in teaching for secure execution and automated testing_ of submissions by students.
[Dodona](https://dodona.ugent.be) is a web platform developed at Ghent University that is used to teach students basic programming skills, and it uses Docker containers to test submissions by students.
This means that both the code testing the students' submissions and the submission itself are executed in a predictable environment, avoiding compatibility issues between the wide variety of configurations used by students.
The containerisation is also used to shield the Dodona servers from bad or even malicious code: memory, time and I/O&nbsp;limits are used to make sure students cannot overload the system.
The web application managing the containers communicates with them by sending configuration information as a JSON document over standard input.
Every Dodona Docker image shares a `main.sh` file that passes through this information to the actual testing framework, while setting up some error handling.
The testing process in the Docker containers sends back the test results by writing a JSON document to its standard output channel.
In June 2019, R support was added to Dodona using an image derived from the `rocker/r-base` image that sets up the `runner` user and and `main.sh` file expected by Dodona\footnote{\href{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}}.
It also installs the packages required for the testing framework and the exercises so that this does not have to happen every time a student's submission is evaluated.
The actual testing of R exercises is done using a custom framework loosely based on \CRANpkg{testthat} \citep{wickham_testthat_2011}.
During the development of the testing framework, it was found that the \pkg{testthat} framework did not provide enough information to its reporter system to send back all the fields required by Dodona to render its feedback.
Right now, multiple statistics courses are developing exercises to automate the feedback for their lab classes.

>>>>

## Teaching
\label{workenvs}

Two use cases demonstrate the practical usefulness and advantages of containerisation in the context of teaching. On the one hand a special case of shared computing environments (see Section&nbsp;\ref{workenvs}), and on the other hand leveraging sandboxing and controlled environments for autograding.

**Prepared environments for teaching** are especially helpful for (a) introductory courses, where students often struggle with the first step of installation and configuration \citep{cetinkaya-rundel_infrastructure_2018}, and (b) courses that require access to a relatively complex setup of software tools, e.g., database systems.
\citet{cetinkaya-rundel_infrastructure_2018} describe how a Docker-based deployment of RStudio (i) avoided problems with troubleshooting individual students' computers and greatly increased engagement through very quickly showing tangible outcomes, e.g., a visualisation, and (ii) reduced demand on teaching and IT staff.
Each student received access to a personal RStudio instance running in a container after authentication with the university login, which gives the benefits of sandboxing and the possibility of limiting resources.
\citet{cetinkaya-rundel_infrastructure_2018} found that for the courses at hand, actual usage of the UI is intermittent so a single cloud-based VM with four cores and 28&nbsp;GB RAM sufficed for over 100 containers.
An example for mitigating _complex setups_ is teaching databases.
R is very useful tool for interfacing with databases, because almost every open source and proprietary database system has an R package that allows users to connect and interact with it.
This flexibility is even broadened by \CRANpkg{DBI} \citep{cran_dbi}, which allows for creating a common API for interfacing these databases, or the \CRANpkg{dbplyr} package \citep{cran_dbplyr}, which runs \CRANpkg{dplyr} \citep{cran_dplyr} code straight against the database as queries.
But learning and teaching these tools comes with the cost of deploying or having access to an environment with the software and drivers installed.
For people teaching R, it can become a barrier if they need to install local versions of database drivers or connect to remote instances which might or might not be made available by IT services.
Giving access to a sandbox for the most common environments for teaching databases is the idea behind [`r-db`](https://github.com/ColinFay/r-db), a Docker image that contains everything needed to connect to a database from R.
Notably, with `r-db`, users do not have to install complex drivers or configure their machine in a specific way.
The `rocker/tidyverse` base image ensures that users can also readily use packages for analysis, display, and reporting.

The idea of a common environment and partitioning allows for using **containers in teaching for secure execution and automated testing** of submissions by students.
First, [Dodona](https://dodona.ugent.be) is a web platform developed at Ghent University that is used to teach students basic programming skills, and it uses Docker containers to test submissions by students.
This means that both the code testing the students' submissions and the submission itself are executed in a predictable environment, avoiding compatibility issues between the wide variety of configurations used by students.
The containerisation is also used to shield the Dodona servers from bad or even malicious code: memory, time and I/O&nbsp;limits are used to make sure students cannot overload the system.
The web application managing the containers communicates with them by sending configuration information as a JSON document over standard input.
Every Dodona Docker image shares a `main.sh` file that passes through this information to the actual testing framework, while setting up some error handling.
The testing process in the Docker containers sends back the test results by writing a JSON document to its standard output channel.
In June 2019, R support was added to Dodona using an image derived from the `rocker/r-base` image that sets up the `runner` user and and `main.sh` file expected by Dodona\footnote{\href{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}}.
It also installs the packages required for the testing framework and the exercises so that this does not have to happen every time a student's submission is evaluated.
The actual testing of R exercises is done using a custom framework loosely based on \CRANpkg{testthat} \citep{wickham_testthat_2011}.
During the development of the testing framework, it was found that the \pkg{testthat} framework did not provide enough information to its reporter system to send back all the fields required by Dodona to render its feedback.
Right now, multiple statistics courses are developing exercises to automate the feedback for their lab classes.

Second, [PrairieLearn](https://github.com/PrairieLearn) is another example of a Docker-based teaching and testing platform.
PrairieLearn is being developed at the University of Illinois at Urbana-Champaign \citep{prairielearn:2018} and has been in extensive use across several faculties along with initial use on some other campuses.
¸It uses Docker containers a key components, both internally for its operations (programmed mainly in Python as well as in Javascript), as well as for two reference containers providing, respectively, Python and R autograders.
A key design decision made by PrairieLearn permits _external_ grading containers to be supplied and accessed via a well-defined interface of invoking, essentially, a single script, `run.sh`.
This script relies on a well-defined file layout containing JSON-based configuration, support files, exam questions, supplementary data, and student submissions.
It returns per-question evaluations as JSON result files, which PrarieLearn evaluates, aggregates and records in a database.
The [Data Science Programming Methods](https://stat430.com) course \citep{stat430:2019} uses this via the custom [`rocker-pl`](https://github.com/stat430dspm/rocker-pl) container \citep{rocker-pl:2019}.\footnote{The reference R container was unavailable at the time, and also relies on a heavier CentOS-based build so that a lighter alternative was established.}
The `rocker-pl` image extends `rocker/r-base` with the \pkg{plr} R package \citep{pkg:plr:2019} for integration into PrarieLearn testing and question evaluation, along with the actual R packages used in instruction and testing for the course in question.
As automated grading of submitted student answers is close to the well-understood problem of unit testing, the \CRANpkg{tinytest} package \citep{CRAN:tinytest} is used for both its core features for testing as well as clean extensibility.
The package \CRANpkg{ttdo} \citep{CRAN:ttdo} utilizes the extensibility of \pkg{tinytest} to display context-sensitive colorized differences between incorrect answers and reference answers using the \CRANpkg{diffobj} package \citep{CRAN:diffobj}.
The setup, described in more detail by \citet{paper:r_autograder}, is an excellent illustration of both the versatility and flexibility offered by Docker-based approaches in teaching and testing.

### 2020-03-04 rockerverse.tex
add teaching use case

\textbf{Teaching} is a further example of where shared computing
environments and sandboxing can greatly improve the process. First,
\emph{prepared environments for teaching} are especially helpful for (a)
introductory courses, where students often struggle with the first step
of installation and configuration
\citep{cetinkaya-rundel_infrastructure_2018}, and (b) courses that
require access to a relatively complex setup of software tools, e.g.,
database systems. \citet{cetinkaya-rundel_infrastructure_2018} describe
how a Docker-based deployment of RStudio (i) avoided problems with
troubleshooting individual students' computers and greatly increased
engagement through very quickly showing tangible outcomes, e.g., a
visualisation, and (ii) reduced demand on teaching and IT staff. Each
student received access to a personal RStudio instance running in a
container after authentication with the university login, which gives
the benefits of sandboxing and the possibility of limiting resources.
\citet{cetinkaya-rundel_infrastructure_2018} found that for the courses
at hand, actual usage of the UI is intermittent so a single cloud-based
VM with four cores and 28~GB RAM sufficed for over 100 containers. An
example for mitigating \emph{complex setups} is teaching databases. R is
very useful tool for interfacing with databases, because almost every
open source and proprietary database system has an R package that allows
users to connect and interact with it. This flexibility is even
broadened by \CRANpkg{DBI} \citep{cran_dbi}, which allows for creating a
common API for interfacing these databases, or the \CRANpkg{dbplyr}
package \citep{cran_dbplyr}, which runs \CRANpkg{dplyr}
\citep{cran_dplyr} code straight against the database as queries. But
learning and teaching these tools comes with the cost of deploying or
having access to an environment with the software and drivers installed.
For people teaching R, it can become a barrier if they need to install
local versions of database drivers or connect to remote instances which
might or might not be made available by IT services. Giving access to a
sandbox for the most common environments for teaching databases is the
idea behind \href{https://github.com/ColinFay/r-db}{\texttt{r-db}}, a
Docker image that contains everything needed to connect to a database
from R. Notably, with \texttt{r-db}, users do not have to install
complex drivers or configure their machine in a specific way. The
\texttt{rocker/tidyverse} base image ensures that users can also readily
use packages for analysis, display, and reporting.

Second, the idea of a common environment and partitioning allows for
using \emph{containers in teaching for secure execution and automated
testing} of submissions by students.
\href{https://dodona.ugent.be}{Dodona} is a web platform developed at
Ghent University that is used to teach students basic programming
skills, and it uses Docker containers to test submissions by students.
This means that both the code testing the students' submissions and the
submission itself are executed in a predictable environment, avoiding
compatibility issues between the wide variety of configurations used by
students. The containerisation is also used to shield the Dodona servers
from bad or even malicious code: memory, time and I/O~limits are used to
make sure students cannot overload the system. The web application
managing the containers communicates with them by sending configuration
information as a JSON document over standard input. Every Dodona Docker
image shares a \texttt{main.sh} file that passes through this
information to the actual testing framework, while setting up some error
handling. The testing process in the Docker containers sends back the
test results by writing a JSON document to its standard output channel.
In June 2019, R support was added to Dodona using an image derived from
the \texttt{rocker/r-base} image that sets up the \texttt{runner} user
and and \texttt{main.sh} file expected by
Dodona\footnote{\href{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}}.
It also installs the packages required for the testing framework and the
exercises so that this does not have to happen every time a student's
submission is evaluated. The actual testing of R exercises is done using
a custom framework loosely based on \CRANpkg{testthat}
\citep{wickham_testthat_2011}. During the development of the testing
framework, it was found that the \pkg{testthat} framework did not
provide enough information to its reporter system to send back all the
fields required by Dodona to render its feedback. Right now, multiple
statistics courses are developing exercises to automate the feedback for
their lab classes.

Another example of a Docker-based teaching and testing platform is
\href{https://github.com/PrairieLearn}{PrairieLearn} which is being
developed at the University of Illinois at Urbana-Champaign
\citep{prairielearn:2018} where it has already been in extensive use
across several faculties (along with initial use on some other
campuses). It uses Docker containers a key components, both internally
for its operations (programmed mainly in Python as well as in
Javascript), as well as for two reference containers providing,
respectively, Python and R autograders. A key design decision made by
PrairieLearn permits \emph{external} grading container to be supplied
and accessed via a well-defined interface of invoking, essentially, a
single script \texttt{run.sh} expecting JSON based configuration,
relying on a well-defined file layout containing support files, exam
questions, supplementary data as well as student submissions, and
returning per-question evaluations as JSON result files which are
evaluated, aggregated and recorded in SQL. The
\href{https://stat430.com}{Data Science Programming Methods} course
\citep{stat430:2019} uses this via the custom
\href{https://github.com/stat430dspm/rocker-pl}{\texttt{rocker-pl}}
container
\citep{rocker-pl:2019}.\footnote{The reference R container was unavailable at the time, and also relies on a heavier CentOS-based build so that a lighter and alternative was established.}
The \texttt{rocker-pl} container extends \texttt{rocker/r-base} with
both the \href{https://github.com/stat430dspm/plr}{\texttt{plr}}
\citep{pkg:plr:2019} custom R package for integration into PrarieLearn
testing and question evaluation, along with the actual R packages used
in instruction and testing for the course in question. As automated
grading of submitted student answers is close to the well-understood
problem of unit testing, the \CRANpkg{tinytest} package
\citep{CRAN:tinytest} is used as it offers both its core features for
testing as well as clean extensibility which package \CRANpkg{ttdo}
\citep{CRAN:ttdo} utilizes to render differences between incorrect
answers and reference answers by invoking the \CRANpkg{diffobj} package
\citep{CRAN:diffobj} for context-sensitive colorized displays of these
differences. The setup, described in more detail by
\citet{paper:r_autograder}, is an excellent illustration of both the
versatility and flexibility offered by Docker-based approaches in
teaching and testing.

@@

The ordering of authors following DN and DE is alphabetical. DN
conceived the article idea,
\href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team},
wrote sections not mentioned below, and revised all sections. DE wrote
the introduction and the section about containerisation and the Rocker
Project, and reviewed all sections. DB wrote the section on
\pkg{outsider}. GD contributed the CARD.com use case. RC \& EH
contributed to the section on interfaces for Docker in R
(\emph{dynverse} and \texttt{dynwrap}). DC contributed content on
Gigantum. ME contributed to the section on processing and deployment to
cloud services. CF wrote paragraphs about \pkg{r-online},
\pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}. LK contributed content on
\pkg{dockr}. SL contributed content on RStudio's usage of Docker. BM
wrote the section on research compendia and made the project Binder
ready. HN \& JN co-wrote the section on the T-Mobile use case. KR wrote
the section about \pkg{holepunch}. NR wrote paragraphs about shared work
environments and GPUs. LS \& NT wrote the section on Bioconductor. PS
wrote the paragraphs about GitLab CI/CD pipelines with Shinyproxy and
OpenFaaS templates. TS \& JW wrote the section on CyVerse. CvP wrote the
section on the usage of Docker containers in Dodona. CW wrote the
sections on Whole Tale and contributed the publication reproducibility
audit use case. NX contributed content on \pkg{liftr}. All authors
approved the final version. This articles was collaboratively written at
\href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The
\href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page},
\href{https://github.com/nuest/rockerverse-paper/commits/master}{commit history},
and
\href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues}
provide a detailed view on the respective contributions.

@@

DN is supported by the project Opening Reproducible Research
(\href{https://www.uni-muenster.de/forschungaz/project/12343}{o2r})
funded by the German Research Foundation (DFG) under project number
\href{https://gepris.dfg.de/gepris/projekt/415851837}{PE~1632/17-1}. The
funders had no role in data collection and analysis, decision to
publish, or preparation of the manuscript. KR was supported in part by a
grant from The Leona M. and Harry B. Helmsley Charitable Trust, award
number 2016PG-BRI004. LS and NT are supported by US NIH / NHGRI awards
U41HG00405 and U24HG010263. CW is supported by the Whole Tale projects
(\url{https://wholetale.org}) funded by the US National Science
Foundation (NSF) under award
\href{https://www.nsf.gov/awardsearch/showAward?AWD_ID=1541450}{OAC-1541450}.
We would like to thank Celeste R. Brennecka from the Science Writing
Support Service of the University of Münster for her editorial support.

>>>>

\hypertarget{teaching}{%
\subsection{Teaching}\label{teaching}}

Two use cases demonstrate the practical usefulness and advantages of
containerisation in the context of teaching. On the one hand a special
case of shared computing environments (see Section~\ref{workenvs}), and
on the other hand leveraging sandboxing and controlled environments for
autograding.

\textbf{Prepared environments for teaching} are especially helpful for
(a) introductory courses, where students often struggle with the first
step of installation and configuration
\citep{cetinkaya-rundel_infrastructure_2018}, and (b) courses that
require access to a relatively complex setup of software tools, e.g.,
database systems. \citet{cetinkaya-rundel_infrastructure_2018} describe
how a Docker-based deployment of RStudio (i) avoided problems with
troubleshooting individual students' computers and greatly increased
engagement through very quickly showing tangible outcomes, e.g., a
visualisation, and (ii) reduced demand on teaching and IT staff. Each
student received access to a personal RStudio instance running in a
container after authentication with the university login, which gives
the benefits of sandboxing and the possibility of limiting resources.
\citet{cetinkaya-rundel_infrastructure_2018} found that for the courses
at hand, actual usage of the UI is intermittent so a single cloud-based
VM with four cores and 28~GB RAM sufficed for over 100 containers. An
example for mitigating \emph{complex setups} is teaching databases. R is
very useful tool for interfacing with databases, because almost every
open source and proprietary database system has an R package that allows
users to connect and interact with it. This flexibility is even
broadened by \CRANpkg{DBI} \citep{cran_dbi}, which allows for creating a
common API for interfacing these databases, or the \CRANpkg{dbplyr}
package \citep{cran_dbplyr}, which runs \CRANpkg{dplyr}
\citep{cran_dplyr} code straight against the database as queries. But
learning and teaching these tools comes with the cost of deploying or
having access to an environment with the software and drivers installed.
For people teaching R, it can become a barrier if they need to install
local versions of database drivers or connect to remote instances which
might or might not be made available by IT services. Giving access to a
sandbox for the most common environments for teaching databases is the
idea behind \href{https://github.com/ColinFay/r-db}{\texttt{r-db}}, a
Docker image that contains everything needed to connect to a database
from R. Notably, with \texttt{r-db}, users do not have to install
complex drivers or configure their machine in a specific way. The
\texttt{rocker/tidyverse} base image ensures that users can also readily
use packages for analysis, display, and reporting.

The idea of a common environment and partitioning allows for using
\textbf{containers in teaching for secure execution and automated
testing} of submissions by students. First,
\href{https://dodona.ugent.be}{Dodona} is a web platform developed at
Ghent University that is used to teach students basic programming
skills, and it uses Docker containers to test submissions by students.
This means that both the code testing the students' submissions and the
submission itself are executed in a predictable environment, avoiding
compatibility issues between the wide variety of configurations used by
students. The containerisation is also used to shield the Dodona servers
from bad or even malicious code: memory, time and I/O~limits are used to
make sure students cannot overload the system. The web application
managing the containers communicates with them by sending configuration
information as a JSON document over standard input. Every Dodona Docker
image shares a \texttt{main.sh} file that passes through this
information to the actual testing framework, while setting up some error
handling. The testing process in the Docker containers sends back the
test results by writing a JSON document to its standard output channel.
In June 2019, R support was added to Dodona using an image derived from
the \texttt{rocker/r-base} image that sets up the \texttt{runner} user
and and \texttt{main.sh} file expected by
Dodona\footnote{\href{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}{https://github.com/dodona-edu/docker-images/blob/master/dodona-r.dockerfile}}.
It also installs the packages required for the testing framework and the
exercises so that this does not have to happen every time a student's
submission is evaluated. The actual testing of R exercises is done using
a custom framework loosely based on \CRANpkg{testthat}
\citep{wickham_testthat_2011}. During the development of the testing
framework, it was found that the \pkg{testthat} framework did not
provide enough information to its reporter system to send back all the
fields required by Dodona to render its feedback. Right now, multiple
statistics courses are developing exercises to automate the feedback for
their lab classes.

Second, \href{https://github.com/PrairieLearn}{PrairieLearn} is another
example of a Docker-based teaching and testing platform. PrairieLearn is
being developed at the University of Illinois at Urbana-Champaign
\citep{prairielearn:2018} and has been in extensive use across several
faculties along with initial use on some other campuses. ¸It uses Docker
containers a key components, both internally for its operations
(programmed mainly in Python as well as in Javascript), as well as for
two reference containers providing, respectively, Python and R
autograders. A key design decision made by PrairieLearn permits
\emph{external} grading containers to be supplied and accessed via a
well-defined interface of invoking, essentially, a single script,
\texttt{run.sh}. This script relies on a well-defined file layout
containing JSON-based configuration, support files, exam questions,
supplementary data, and student submissions. It returns per-question
evaluations as JSON result files, which PrarieLearn evaluates,
aggregates and records in a database. The
\href{https://stat430.com}{Data Science Programming Methods} course
\citep{stat430:2019} uses this via the custom
\href{https://github.com/stat430dspm/rocker-pl}{\texttt{rocker-pl}}
container
\citep{rocker-pl:2019}.\footnote{The reference R container was unavailable at the time, and also relies on a heavier CentOS-based build so that a lighter alternative was established.}
The \texttt{rocker-pl} image extends \texttt{rocker/r-base} with the
\pkg{plr} R package \citep{pkg:plr:2019} for integration into
PrarieLearn testing and question evaluation, along with the actual R
packages used in instruction and testing for the course in question. As
automated grading of submitted student answers is close to the
well-understood problem of unit testing, the \CRANpkg{tinytest} package
\citep{CRAN:tinytest} is used for both its core features for testing as
well as clean extensibility. The package \CRANpkg{ttdo}
\citep{CRAN:ttdo} utilizes the extensibility of \pkg{tinytest} to
display context-sensitive colorized differences between incorrect
answers and reference answers using the \CRANpkg{diffobj} package
\citep{CRAN:diffobj}. The setup, described in more detail by
\citet{paper:r_autograder}, is an excellent illustration of both the
versatility and flexibility offered by Docker-based approaches in
teaching and testing.

@@

The ordering of authors following DN and DE is alphabetical. DN
conceived the article idea,
\href{https://github.com/nuest/rockerverse-paper/issues/3}{initialised the formation of the writing team},
wrote sections not mentioned below, and revised all sections. DE wrote
the introduction and the section about containerisation and the Rocker
Project, and reviewed all sections. DB wrote the section on
\pkg{outsider}. GD contributed the CARD.com use case. RC \& EH
contributed to the section on interfaces for Docker in R
(\emph{dynverse} and \texttt{dynwrap}). DC contributed content on
Gigantum. ME contributed to the section on processing and deployment to
cloud services. CF wrote paragraphs about \pkg{r-online},
\pkg{dockerfiler}, \pkg{r-ci} and \pkg{r-db}. LK contributed content on
\pkg{dockr}. SL contributed content on RStudio's usage of Docker. BM
wrote the section on research compendia and made the project Binder
ready. HN \& JN co-wrote the section on the T-Mobile use case. KR wrote
the section about \pkg{holepunch}. NR wrote paragraphs about shared work
environments and GPUs. LS \& NT wrote the section on Bioconductor. PS
wrote the paragraphs about GitLab CI/CD pipelines with Shinyproxy and
OpenFaaS templates. TS \& JW wrote the section on CyVerse. CvP wrote the
section on the usage of Docker containers in Dodona. CW wrote the
sections on Whole Tale and contributed content about publication
reproducibility audits. NX contributed content on \pkg{liftr}. All
authors approved the final version. This articles was collaboratively
written at
\href{https://github.com/nuest/rockerverse-paper/}{https://github.com/nuest/rockerverse-paper/}.
The
\href{https://github.com/nuest/rockerverse-paper/graphs/contributors}{contributors page},
\href{https://github.com/nuest/rockerverse-paper/commits/master}{commit history},
and
\href{https://github.com/nuest/rockerverse-paper/issues/}{discussion issues}
provide a detailed view on the respective contributions.

@@

DN is supported by the project Opening Reproducible Research
(\href{https://www.uni-muenster.de/forschungaz/project/12343}{o2r})
funded by the German Research Foundation (DFG) under project number
\href{https://gepris.dfg.de/gepris/projekt/415851837}{PE~1632/17-1}. The
funders had no role in data collection and analysis, decision to
publish, or preparation of the manuscript. KR was supported in part by a
grant from The Leona M. and Harry B. Helmsley Charitable Trust, award
number 2016PG-BRI004. LS and NT are supported by US NIH / NHGRI awards
U41HG00405 and U24HG010263. CW is supported by the Whole Tale project
(\url{https://wholetale.org}) funded by the US National Science
Foundation (NSF) under award
\href{https://www.nsf.gov/awardsearch/showAward?AWD_ID=1541450}{OAC-1541450}.
We would like to thank Celeste R. Brennecka from the Science Writing
Support Service of the University of Münster for her editorial support.
